Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-08-26 main.tex
Updates from Overleaf

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
% my added packages and commands =======================
\usepackage[numbers]{natbib} %removed cite package
\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{caption}

@@

\begin{table*}[t]
\centering
\caption{Summary of select works in sound-matching. Gen. refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
\scriptsize % Reduce font size
\resizebox{\textwidth}{!}{
\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
\hline
\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
\hline
Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
\hline
Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-diff FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
\hline
Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-diff VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
\hline
Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
\hline
Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
\hline
Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & Non-diff & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
\hline
Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
\hline
Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
\hline
Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Gradient desc. & Replicate & In & 2023 & Iter \\
\hline
Barkan \textit{et al.}~\cite{barkan2023inversynth}& Diff. VST Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
\\
\hline
Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
\hline
Cherep \textit{et al.}~\cite{creativecherep2024} & DDSP & CLAP & Evolutionary & Imitate & Out & 2024 & Iter \\
\hline
\end{tabular}
}

>>>>

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
% my added packages and commands =======================
\usepackage[numbers]{natbib} %removed cite package
\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{caption}

@@

\begin{table*}[t]
\centering
\caption{Summary of select works in sound-matching. Gen. refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
\scriptsize % Reduce font size
\resizebox{\textwidth}{!}{
\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
\hline
\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
\hline
Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-differentiable FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
\hline
Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-differentiable FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
\hline
Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-differentiable VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
\hline
Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-differentiable VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
\hline
Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
\hline
Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & DSP & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
\hline
Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
\hline
Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
\hline
Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Gradient desc. & Replicate & In & 2023 & Iter \\
\hline
Barkan \textit{et al.}~\cite{barkan2023inversynthII}& Diff. Synthesizer Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
\\
\hline
Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
\hline
Cherep \textit{et al.}~\cite{creativecherep2024} & DDSP & CLAP & Evolutionary & Imitate & Out & 2024 & Iter \\
\hline
\end{tabular}
}

### 2025-08-26 revised_main.tex
Updates from Overleaf

\section{Background and Past Works}
\label{sec:background}
This section provides an introduction to digital audio synthesis, audio similarity measures, and a formal definition of sound-matching. We then provide a typology of sound-matching tasks using previous works as well as conceptual approaches that may be promising but though underexplored. Finally, a discussion on the current gaps of knowledge in the field is provided.

@@

In addition, Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
Finetuning (ITF). For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}. Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF does not improve IS2 in spectral and manual performance metrics.

@@

% masuda and esling both noted out-of-domain drop in performance
% yee king noted performance drop in manual surveys 
%p-loss bad : masuda-improving, han2023, esling-flowsynth

>>>>

% need to simplify the intro, and expand the hypothesis
\section{Introduction}
Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}. Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation. However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds. One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods. As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.

@@

Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
Finetuning (ITF). For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}. Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF to IS2 generally worsened the spectral and manual performance metrics.

@@

Some recent works have opted for the use of audio embeddings as a similarity metric. In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects. In their approach, a desired sound-effect is described in text and embedded using clap, and gradient free optimizer~\cite{evosax2022github} iteratively updates the synthesizer's parameters to minimize difference in the text and output sound embeddings. Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}, however, it did yield better scores for ``artistic-interpretation'' (or what we refer to as imitation).

% masuda and esling both noted out-of-domain drop in performance
% yee king noted performance drop in manual surveys 
%p-loss bad : masuda-improving, han2023, esling-flowsynth
% manual performance measure: barkan, masuda2021soundmatch, yee-king
